Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-10-28 files/Chapter4/ModalAnalysisTimo.tex
chapter 4

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
An example of modal analysis on a Timoshenko beam is given in section \ref{sec:existence:ModalAnalysis}. In this chapter, more general theory for modal analysis of te Timoshenko beam theory is discussed, as well as some more examples of the application of the theory.

@@

\textcolor{red}{*****************SUGGESTION****************************}

\subsubsection*{Stragety for determining the eigenvalues and eigenvectors}

@@

\textcolor{red}{*****************OLD alles onder moontlik weg vat****************************}

The next step is to determine the solution of the general eigenvalue problem. The strategy to obtain the eigenvalues and eigenvectors are described in detail in \cite{VV06} using an example. A summary of the strategy is presented here.

The boundary conditions at $x=1$ is then substituted which results in a homogeneous system of linear equations of the form
\begin{eqnarray*}
A \bar{b} = \bar{0}
\end{eqnarray*}
This system either has the zero

 This system has either the zero solution or infinitely many solutions. To ensure a non-zero solution, the determinant of the coefficient matrix $A$ of the system of equations is set equal to 0. The equation $\textrm{det}(A) = 0$ is called the frequency equation. The frequency equation has infinitely many solutions and each solution corresponds to an eigenvalue.

 For each eigenvalue, a unique solution for the pair $\bar{b}$ can be obtained. A eigenfunction for $u$ and $\phi$ is obtained using $\bar{b}$ and corresponds to the relevant eigenvalue.

The following two sections provide examples of this strategy applied to beam models. These examples are also in \cite{VV06}.

**sluit af***
******NB Verduidelik $\alpha$
\end{document}
%obvisouly the solution space is 4 dimensional . bc at 0 makes it at most two.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
An example of modal analysis on a Timoshenko beam is given in section \ref{sec:existence:ModalAnalysis}. In this chapter, more general theory for modal analysis of te Timoshenko beam theory is discussed, as well as more examples of the application of the theory.

@@

\subsubsection*{Strategy for determining the eigenvalues and eigenvectors}

@@

\end{document}
%obvisouly the solution space is 4 dimensional . bc at 0 makes it at most two.

### 2023-10-28 files/Chapter4/SP06.tex
chapter2
chapter 4

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{SP06}
Consider a study by Stephen and Puchegger in 2006, article [SP06]. This study investigated the validity of the Timoshenko beam theory by hand of empirical and numerical data. The approach of the study is to compare the natural frequencies of a Timoshenko beam, to that of a three-dimensional elastic beam. The authors conducted comparisons between the models using theoretical methods as well as results from an empirical study on a physical beam, conducted by the authors.\\

The authors decided on a free-free beam configuration, as it is the easiest to achieve in the physical experiment. The natural frequencies of the Timoshenko beam theory was obtained by using frequency equations from an article by Levison and Cooke [LC81]. In theoretical approach of the three-dimensional beam, two methods were considered. A commercial finite elements method (ANSYS), and a resonant ultrasound spectroscopy (RUS) technique.\\

This section discusses the results of [SP06].

\subsection{Mathematical models}
Let $\Omega$ be the reference configuration of a free-free beam with square cross-section. The authors of [SP06] do not formulate the models in the article. Consider Problem T-4 from section \ref{ssec:1D_Model:ModelProblems} for the free-free Timoshenko beam and Problem 3D-2 from section \ref{ssec:3D_Model:ModelProblems} for the free-free three-dimensional beam. The model problems are given here again for the reader's convenience.

\subsubsection*{Problem T-4}
Find a function $u$, satisfying the equations of motion \eqref{eq:1D_Model:EquationOfMotion1D}- \eqref{eq:1D_Model:EquationOfMotion2D} and constitutive equations \eqref{eq:1D_Model:ConstitutiveEquations1D}- \eqref{eq:1D_Model:ConstitutiveEquations2D}.\\

@@

\subsection{Experimental setup}
Next the experiment conducted by [SP06] is discussed. A short, aluminium alloy beam with near square cross-section is suspended at both ends. The beam is suspended by carbon fibre loops.

@@

To vibrate the beam, the authors of [SP06] excited the carbon fibre loops and the frequencies were measured using two piezoelectric transducers. During the vibration, the beam will be momentarily free at both endpoints. The natural frequencies were obtained by sweeping through frequencies until a resonance was found.\\

The measured parameters of the beam is given in [SP06]. Since the physical beam is only nearly square, the plane has two distinct planes. The plane of the beam with a larger diameter is referred to as the stiff plane and the plane of the beam with smaller diameter is referred to as the flexible plane, by [SP06].\\

For the flexible plane $\alpha = \pm 190$ and for the stiff plane $\alpha = \pm 189$. The parameter $\gamma = 0.2830$. The natural frequencies $f_k$ is defined as 
\begin{eqnarray*}
	f_k = \sqrt{\frac{\lambda_k}{2\pi}}
\end{eqnarray*} with $\lambda_k$ the $k$'th eigenvalue. These natural frequencies $f_k$ are dimensionless, and can be scaled back by multiplying the natural frequency $f_k$ by $t_0=\ell\sqrt{\frac{\rho}{G\kappa^2}}$, as defined in Chapter 1.\\

The cut-off frequency given in [SP06] can be expressed as
\begin{eqnarray*}
	\omega_{co} = \sqrt{\frac{\kappa^2 A G}{\rho I}} = \frac{\alpha}{t_0}
\end{eqnarray*}
with the dimensionless cut-off frequency $\alpha$.

\subsection{Results from SP06}
The following table contains relevant results are obtained by [SP06]. 
% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
	\makebox[\textwidth]{
	\caption{Results from [SP06] (excluding RUS).}
	\begin{tabular}{||c|c|cc|cc|c||}
		\hline
		{n} & {Measured} & {3D FEM} & {\% Error} & {Timoshenko Beam} & {\% Error} & Side \\
		\hline
		1     & 27359.6 & 27417.6 & 0,21\% & 27407.1 & 0,17\% & Flexible \\
		  	  & 27423.8 & 27515.6 & 0,33\% & 27505.3 & 0,30\% & Stiff \\
		2     & 60862   & 60882.0 & 0,03\% & 60851.1 & -0,02\% & Flexible \\
		 	  & 61098.3 & 61022.0 & -0,12\% & 60992.1 & -0,17\% & Stiff \\
		3     & 97609.5 & 97734.4 & 0,13\% & 97796.0 & 0,19\% & Flexible \\
			  & 97852.4 & 97881.5 & 0,03\% & 97945.4 & 0,09\% & Stiff \\
		4     & 161494 & 131658 & 0,12\% & 132277 & 0,60\% & Flexible \\
			  & 131732 & 131675 & -0,04\% & 132308 & 0,44\% & Stiff \\
		5     & 161352 & 161390 & 0,02\% & 163547 & 1,36\% & Stiff \\
			  & 161538 & 161517 & -0,01\% & 163611 & 1,31\% & Flexible \\
		6     & 165183 & 164887 & -0,18\% & 169108 & 2,38\% & Stiff \\
			  & 165598 & 165398 & -0,12\% & 169634 & 2,44\% & Flexible \\
		7     & 194863 & 194933 & 0,04\% & 202352 & 3,84\% & Flexible \\
		   	  & 194973 & 195032 & 0,03\% & 202115 & 3,66\% & Stiff \\
		8     & 195869 & 195977 & 0,06\% & 203319 & 3,80\% & Stiff \\
			  & 195908 & 196097 & 0,10\% & 203518 & 3,88\% & Flexible \\
		9     & 213501 &       &       & 241202 & 12,97\% & Flexible \\
			  & 213635 &       &       & 241067 & 12,84\% & Stiff \\
  {10 and 11} & 220556 &       &       & 247954 & 12,42\% & Flexible \\
			  & 220702 &       &       & 281542 & 27,57\% & Flexible \\
			  & 221010 &       &       & 247782 & 12,11\% & Stiff \\
			  & 221092 &       &       & 281408 & 27,28\% & Stiff \\
		\hline
	\end{tabular}%
	\label{tab:addlabel}%
}
\end{table}%

@@

Table 4.3 shows that the Timoshenko model compares well to the measured results from the experiment. The first 8 natural frequencies for the stiff and flexible planes are very close to the natural frequencies of the three-dimensional and physical beam.\\

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Empirical and numerical examination of a Timoshenko beam} \label{sec:SP06}
Consider a study by Stephen and Puchegger in 2006, article \cite{SP06}. This study investigated the validity of the Timoshenko beam theory by hand of empirical and numerical data. The approach of the study is to compare the natural frequencies of a Timoshenko beam, to that of a three-dimensional elastic beam. The authors conducted comparisons between the models using theoretical methods as well as results from an empirical study on a physical beam, conducted by the authors.

The authors decided on a free-free beam configuration. The natural frequencies of the Timoshenko beam theory was obtained by using frequency equations from an article by Levison and Cooke \cite{LC81}. In theoretical approach of the three-dimensional beam, two methods were considered. A commercial finite elements method (ANSYS), and a resonant ultrasound spectroscopy (RUS) technique.

This section discusses the results of \cite{SP06}.

\subsection{Mathematical models}
Let $\Omega$ be the reference configuration of a free-free beam with square cross-section. The authors of \cite{SP06} do not formulate the models in the article. Consider Problem T-4 from section \ref{ssec:1D_Model:ModelProblems} for the free-free Timoshenko beam and Problem 3D-2 from section \ref{ssec:3D_Model:ModelProblems} for the free-free three-dimensional beam. The boundary conditions and reference configurations of the model problems are given below.

\subsubsection*{Problem T-4}
Find a function $u$, satisfying the equations of motion \eqref{eq:1D_Model:EquationOfMotion1D}- \eqref{eq:1D_Model:EquationOfMotion2D} and constitutive equations \eqref{eq:1D_Model:ConstitutiveEquations1D}- \eqref{eq:1D_Model:ConstitutiveEquations2D}.

@@

\subsection{Experimental setup}
Next the experiment conducted by \cite{SP06} is discussed. A short, aluminium alloy beam with near square cross-section is suspended at both ends. The beam is suspended by carbon fibre loops.

@@

To vibrate the beam, the authors of \cite{SP06} excited the carbon fibre loops and the frequencies were measured using two piezoelectric transducers. During the vibration, the beam will be momentarily free at both endpoints. The natural frequencies were obtained by sweeping through frequencies until a resonance was found.

The measured parameters of the beam is given in \cite{SP06}. Since the physical beam is only nearly square, the plane has two distinct planes. The plane of the beam with a larger diameter is referred to as the stiff plane and the plane of the beam with smaller diameter is referred to as the flexible plane, by \cite{SP06}.

For the flexible plane $\alpha = \pm 190$ and for the stiff plane $\alpha = \pm 189$. The parameter $\gamma = 0.2830$. The natural frequencies $f_k$ is defined as 
\begin{eqnarray*}
	f_k = \sqrt{\frac{\lambda_k}{2\pi}}
\end{eqnarray*} with $\lambda_k$ the $k$'th eigenvalue. These natural frequencies $f_k$ are dimensionless, and can be scaled back by multiplying the natural frequency $f_k$ by $t_0=\ell\sqrt{\frac{\rho}{G\kappa^2}}$, as defined in Chapter 1.

The cut-off frequency given in \cite{SP06} can be expressed as
\begin{eqnarray*}
	\omega_{co} = \sqrt{\frac{\kappa^2 A G}{\rho I}} = \frac{\alpha}{t_0}
\end{eqnarray*}
with the dimensionless cut-off frequency $\alpha$.

\subsection{Results from SP06}
The following table contains relevant results are obtained by \cite{SP06}. 
% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
	\makebox[\textwidth]{
	\caption{Results from \cite{SP06} (excluding RUS).}
	\begin{tabular}{||c|c|cc|cc|c||}
		\hline
		{n} & {Measured} & {3D FEM} & {\% Error} & {Timoshenko Beam} & {\% Error} & Side \\
		\hline
		1     & 27359.6 & 27417.6 & 0,21\% & 27407.1 & 0,17\% & Flexible \\
		  	  & 27423.8 & 27515.6 & 0,33\% & 27505.3 & 0,30\% & Stiff \\
		2     & 60862   & 60882.0 & 0,03\% & 60851.1 & -0,02\% & Flexible \\
		 	  & 61098.3 & 61022.0 & -0,12\% & 60992.1 & -0,17\% & Stiff \\
		3     & 97609.5 & 97734.4 & 0,13\% & 97796.0 & 0,19\% & Flexible \\
			  & 97852.4 & 97881.5 & 0,03\% & 97945.4 & 0,09\% & Stiff \\
		4     & 161494 & 131658 & 0,12\% & 132277 & 0,60\% & Flexible \\
			  & 131732 & 131675 & -0,04\% & 132308 & 0,44\% & Stiff \\
		5     & 161352 & 161390 & 0,02\% & 163547 & 1,36\% & Stiff \\
			  & 161538 & 161517 & -0,01\% & 163611 & 1,31\% & Flexible \\
		6     & 165183 & 164887 & -0,18\% & 169108 & 2,38\% & Stiff \\
			  & 165598 & 165398 & -0,12\% & 169634 & 2,44\% & Flexible \\
		7     & 194863 & 194933 & 0,04\% & 202352 & 3,84\% & Flexible \\
		   	  & 194973 & 195032 & 0,03\% & 202115 & 3,66\% & Stiff \\
		8     & 195869 & 195977 & 0,06\% & 203319 & 3,80\% & Stiff \\
			  & 195908 & 196097 & 0,10\% & 203518 & 3,88\% & Flexible \\
		9     & 213501 &       &       & 241202 & 12,97\% & Flexible \\
			  & 213635 &       &       & 241067 & 12,84\% & Stiff \\
  {10 and 11} & 220556 &       &       & 247954 & 12,42\% & Flexible \\
			  & 220702 &       &       & 281542 & 27,57\% & Flexible \\
			  & 221010 &       &       & 247782 & 12,11\% & Stiff \\
			  & 221092 &       &       & 281408 & 27,28\% & Stiff \\
		\hline
	\end{tabular}%
	\label{tab:addlabel}%
}
\end{table}%

@@

Table 4.3 shows that the Timoshenko model compares well to the measured results from the experiment. The first 8 natural frequencies for the stiff and flexible planes are very close to the natural frequencies of the three-dimensional and physical beam.
